### 2019-04-30 report/paper.tex
report

\documentclass[journal, a4paper]{IEEEtran}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
% some very useful LaTeX packages include:

%\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath,eqparbox,booktabs,xparse}
\usepackage{smartdiagram}
\usetikzlibrary{quotes}
%\usepackage{psfrag}   
%\usepackage{subfigure}
\usepackage{neuralnetwork}
\usepackage{url}        % Written by Donald Arseneau
%\usepackage{stfloats}  % Written by Sigitas Tolusis
\usepackage{xcolor}
%\interdisplaylinepenalty=2500
\usepackage{listings}

\lstset{basicstyle=\ttfamily, keywordstyle=\bfseries}
\usetikzlibrary{arrows.meta, positioning}
\tikzset{%
    block/.style = {draw=gray, rectangle, thick,
        minimum height=6em, text width=4em, align=center}
}

\makeatletter
\NewDocumentCommand{\eqmathbox}{o O{c} m}{%
  \IfValueTF{#1}
    {\def\eqmathbox@##1##2{\eqmakebox[#1][#2]{$##1##2$}}}
    {\def\eqmathbox@##1##2{\eqmakebox{$##1##2$}}}
  \mathpalette\eqmathbox@{#3}
}
% Your document starts here!
\begin{document}

% Define document title and author
	\title{Project 1 of Deep Learning\\ \textit{\Large{Classification, weight sharing, auxiliary losses}}}
	\author{Francis \textbf{Damachi} - Costanza \textbf{Volpini}}
	\markboth{Project 1 - Deep Learning EE559 - EPFL}{}
	\maketitle

\begin{abstract}
The aim of this project was to show the impact of weight sharing and the use of auxiliary loss. The task was to compare two digits in a two-channel image. Starting from linear model until some more complex non-linear model, we have found out that convolutional neural networks represents the best solution to classify images.
\end{abstract}

\section{Introduction}
\label{sec:intro}
\textit{Convolutional neural network} (CNN) represents the most powerful model to analyse images. Indeed, CNN uses the pixel locality to learn better.

We have started the project training simple linear models, in particular \textit{Linear Regression} and \textit{Logistic Regression} (see Section \ref{sec:linearmodel}) but we have notice that these models are too simple and not consider the pixel locality. We have then used a more complex model \textit{Neural Network} with one-loss and two-losses (as explained and showed in Section \ref{sec:nnmodel}), we got a bad accuracy caused by the fact that we have a model composed by just linear layer. The last model that we have implemented is the \textit{Convolutional neural network}, with this model we got a very good approximation (see Section \ref{sec:cnnmodel}).
In all the models, we have used \textit{Adam's method} instead of \textit{stochastic gradient descent} as optimizer since it use an adaptive learning rate (the learning rate of SGD has an equivalent type of effect for all the weights/parameters of the model \cite{reference0}).
As criterion, we have used \textit{Mean Square Error} for all the models.

Section \ref{sec:data} explain the structure of the dataset, Section \ref{sec:codestruc} describes briefly the content of each file. In Section \ref{sec:comparison}, \ref{sec:conclusion} we have showed the performance accuracy obtained by each model with a fixed seed, explaining why we got these results.

\section{Data}
\label{sec:data}
The data set consists of an input series of $2\times14\times14$ tensor (2 images in grayscale). Training and test set is $1000$ pairs each. In total we will have six tensors: images ($N\times2\times14\times14$), prediction (boolean), classes of the two digits ($N\times2$), both for training and test sets. The MNIST database is composed by handwritten digits (from $0$ to $9$); to load the dataset we have used the provided function \texttt{generate\_pair\_sets(N)}.

\section{Code structure}
\label{sec:codestruc}
In this section contains the description of each file.
\begin{itemize}
    \item \texttt{scripts/data.py}: class to generate the dataset. Contains different methods (e.g. flat the input, get a dataset in 2D or in 3D, enable the hot-encoding).
    \item \texttt{scripts/model.py}: general class to define a model to train and test it (with corresponding plot and history).
    \item \texttt{scripts/models\_implemented.py}: contains all the model classes (e.g. NNModel1Loss, CNNModel2Loss).
    \item \texttt{main.py}: contains examples of each models, in order to call and train a model.
\end{itemize}

\section{Linear Models}
\label{sec:linearmodel}
We have implemented \textit{Linear Regression} and \textit{Logistic Regression}. Regression is represented by a final layer that it is linear. In the logistic regression we have used a sigmoid as an activation function in the output layer. A linear layer accept as input a vector of values, that is the reason why we need to flat the input (the raw input is composed by 2D images). In fact a linear layer give a weight to each pixel, then it will accept a vector composed by N rows and D dimensions (columns), in our case it will have size $14 \times 14 \times 2$.

% CNN sfrutta localita pixel e le tratta come img - no flatten.

% ASCOLTARE DA AUDIO 19 Aprile

%TODO: mettere size flatted come input???
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.4\textwidth]{linearregression.png}
    \caption{Linear Regression model}
    \label{fig:linearregression}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.4\textwidth]{logistic.png}
    \caption{Logistic Regression model}
    \label{fig:logisticregression}
\end{figure}

\section{Neural Network Models}
\label{sec:nnmodel}

\section{Convolutional Neural Network Models}
\label{sec:cnnmodel}

\section{Comparison of models}
\label{sec:comparison}

\section{Conclusion}
\label{sec:conclusion}
% %%%%%%%%%%%%%%%%%%%%%

% Now we need a bibliography:
\begin{thebibliography}{999}

	%Each item starts with a \bibitem{reference} command and the details thereafter.
	\bibitem{reference0}
    	optim.Adam vs optim.SGD. Let’s dive in.
    	\url{https://medium.com/@Biboswan98/optim-adam-vs-optim-sgd-lets-dive-in-8dbf1890fbdc}

% 	\bibitem{reference1}
% 	    Neural networks and deep learning, Michael Nielsen.

%     \bibitem{reference2}
%     Notes of course Deep Learning (EE-559), Prof. François Fleuret.

\end{thebibliography}

% Your document ends here!
\end{document}

### 2019-04-30 report/paper.tex
report wip

@@
 
+For all the models we have explicitely separated in the code the feature extractor, used to process the input and extract high level features, from classifier(s), a simple linear layer that we have used to classify.
+
 Section \ref{sec:data} explain the structure of the dataset, Section \ref{sec:codestruc} describes briefly the content of each file.
@@
 In fact a linear layer give a weight to each pixel, then it will accept a vector composed by N rows and D dimensions (columns), in our case it will have size $14 \times 14 \times 2$.
+Linear layers bad generalize images, since it just memorize the train images, that causes a good train accuracy but a very low test accuracy.
 
-% CNN sfrutta localita pixel e le tratta come img - no flatten.
-
-% ASCOLTARE DA AUDIO 19 Aprile
+During the train of linear models we have noticed that sometimes the accuracy and loss could both decreases, that can happen if we have a case like the blue line in Fig. \ref{fig:lossacc}. 
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.4\textwidth]{lossacc.png}
+    \caption{Red line has an higher loss but a better accuracy.
+Instead, blue line has a lower loss but it predicts bad (lower accuracy).
+We can conclude saying that blue line fits well samples that already known.}
+    \label{fig:lossacc}
+\end{figure}
+Linear Regression has a strange behavior due to the fact that it has not have a sigmoid and the loss MSE works very bad in a classification task without mapping the value in a range.
 
@@
 \label{sec:nnmodel}
+We have implemented two models for the neural network, the first one with 1 loss and the second one with 2 losses, as showed in Fig.
+\ref{fig:nn1} and Fig.
+\ref{fig:nn2}.
+Both models use the same feature extractors but they differs for the classifier.
+The first one has a boolean classifier that returns 1 if the first image is smaller then the second one.
+The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images (we have used an hot encoding to give a target to each node).
+%check se img1 < img2 o se era l'inverso!
+For these two models the images inputs are passed as 2D images.
+Since the feature extractor processes both images together (it does not consider the two images independently), we can use a digit classifier with $20$ nodes as output ($10$ nodes for each images; i.e.
+from $0$ to $9$ as value of digit).
+We were expected to have a better result using 2 losses; but since it is a model with just linear layers we could have improved the computation considering both the images independently; we have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
+We have observed that two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
+
+% Aggiungere un preprocessing step (flatten nel caso delle reti basate sui layer lineari e niente nel caso della rete conv3d)
 
@@
 \label{sec:cnnmodel}
+Two models of convolutional neural networks were trained, both received as input 3D images since we have decided to use a \textit{conv3d}.
+Conv3d are powerful cnn filter that are able to traverse the image in \textit{x}, \textit{y} and \textit{z} axes; that make possible to process both the two images in a independent way.
+CNN are powerfull since they use pixel locality, then instead of flatting the input as done in previous model we can just consider the 3D raw images.
+As done, in Sec.
+\ref{sec:nnmodel} we have decided to compare two models, one with one loss and the other one with two losses, as done with Neural Network models.
+%AGGIUNGERE COMPARAZIONE PER DIRE QUALE DEI DUE MODELLI É STATO MEGLIO
+We have used the dropout to generalize better and then obtain a more robust model.
 
@@
 \label{sec:comparison}
+Since linear models are really sensible to the weights setting 
+we have decided to use a unique seed to compare the models.
+
+% #TODO: (REPORT) for 1 to 10 of seed and to plotbox per far vedere outliers --> pensare al codice
+% #TODO: (REPORT) fare tabella con num\_params per ogni modello
 
@@
 \label{sec:conclusion}
+As future improvement we would like to implement the cross validation.
+%TODO: add
 % %%%%%%%%%%%%%%%%%%%%%

### 2019-04-30 main.py

@@ -8,9 +8,6 @@ model = 4
 
 d = Data()
 
-#TODO: (REPORT) for 1 to 10 of seed and to plotbox per far vedere outliers
-#TODO: (REPORT) fare tabella con num_params per ogni modello
-
 train_input = None
 train_target = None
 train_classes = None
@@ -27,8 +24,8 @@ else: # append img 1 and 2
 
 #########################################################
 
-#TODO: cross validation to find best parameters
-# Adam’s method considered as a method of Stochastic Optimization is a technique implementing adaptive learning rate. Whereas in normal SGD the learning rate has an equivalent type of effect for all the weights/parameters of the model.
+# Adam’s method considered as a method of Stochastic Optimization is a technique implementing adaptive learning rate.
+# Whereas in normal SGD the learning rate has an equivalent type of effect for all the weights/parameters of the model.
 
 ###################### 1. LINEAR MODEL #####################
 if(model == 1):

### 2019-04-30 scripts/models_implemented.py

@@ -213,7 +213,7 @@ class CNNModel1Loss(Model):
             nn.Linear(8 * 2 * 7 * 7, 32),
             nn.ReLU(),
             nn.Linear(32, output_size),
-            nn.Sigmoid() # or sigmoid (TODO: attenzione al target durante il train con la sigmoid)
+            nn.Sigmoid()
         )
 
         self.optimizer = optimizer(self.parameters())

### 2019-04-30 main.py
remove learning rate from old models + report wip

@@ -2,7 +2,7 @@ import torch
 from scripts.data import Data
 from scripts.models_implemented import *
 
-model = 4
+model = 3
 
 ################### GENERATE DATASETS ###################
 
@@ -29,10 +29,14 @@ else: # append img 1 and 2
 
 ###################### 1. LINEAR MODEL #####################
 if(model == 1):
-    torch.manual_seed(1)
+    # torch.manual_seed(1)
 
     # Train the model
     model_linear = LinearRegression()
+
+    print("Number of parameters: {}".format(model_linear.number_params()))
+    print("Number of parameters of feature_extractor: /")
+    print("Number of parameters of classifier: {}".format(model_linear.number_params(model_linear.classifier)))
 
     model_linear.fit(
         train_input, train_target,
@@ -47,8 +51,12 @@ if(model == 1):
 #################### 2. LOGISTIC MODEL #####################
 elif(model == 2):
     # Train the model
-    torch.manual_seed(1)
+    # torch.manual_seed(1)
     model_logistic = LogisticRegression()
+
+    print("Number of parameters: {}".format(model_logistic.number_params()))
+    print("Number of parameters of feature_extractor: /")
+    print("Number of parameters of classifier: {}".format(model_logistic.number_params(model_logistic.classifier)))
 
     model_logistic.fit(
         train_input, train_target,

### 2019-04-30 scripts/models_implemented.py

@@ -14,8 +14,7 @@ class LinearRegression(Model):
             'features_in': features_in,
             'features_out': features_out,
             'optimizer': optimizer,
-            'criterion': criterion,
-            'learning_rate': learning_rate
+            'criterion': criterion
         }
         self.classifier = nn.Sequential(
             nn.Linear(features_in, features_out)
@@ -37,8 +36,7 @@ class LogisticRegression(Model):
             'features_in': features_in,
             'features_out': features_out,
             'optimizer': optimizer,
-            'criterion': criterion,
-            'learning_rate': learning_rate
+            'criterion': criterion
         }
         self.classifier = nn.Sequential(
             nn.Linear(features_in, features_out)

### 2019-05-02 cross_validation_report.py
important changes + cross validation to generate plotbox report

@@ -1,0 +1,67 @@
+from sklearn.model_selection import KFold
+from scripts.data import Data
+from scripts.models_implemented import *
+import matplotlib.pylab as plt
+
+# get the data
+d = Data()
+
+# define the number of splits for the cross validation
+n_splits=5
+
+# define the parameters for the fit of the models
+model_class_fit_params = {
+    LinearRegression: {'epochs': 1000, 'batch_size': 1000},
+    LogisticRegression: {'epochs': 1000, 'batch_size': 1000},
+    NNModel1Loss: {'epochs': 500, 'batch_size': 1000},
+    NNModel2Loss: {'epochs': 1000, 'batch_size': 1000},
+    CNNModel1Loss: {'epochs': 70, 'batch_size': 50},
+    CNNModel2Loss: {'epochs': 70, 'batch_size': 50},
+}
+
+# cross validate
+kf = KFold(n_splits=n_splits)
+accuracies = {}
+for model_class, fit_params in model_class_fit_params.items():
+    model_class_name = model_class.__name__
+    print('Evaluating', model_class_name)
+    train_input, train_target, test_input, test_target = model_class.reshape_data(d)
+
+    accuracies[model_class_name] = {'cross_val_accuracies': [], 'train_accuracy': None, 'test_accuracy': None}
+    # cross validdation estimate the score we will obtain with the test set
+    for train_index, test_index in kf.split(train_input):
+        X_train, X_test = train_input[train_index], train_input[test_index]
+        if isinstance(train_target, tuple):
+            y_train = tuple([y[train_index] for y in train_target])
+            y_test = tuple([y[test_index] for y in train_target])
+        else:
+            y_train, y_test = train_target[train_index], train_target[test_index]
+
+        model = model_class()
+        model.fit(
+            X_train, y_train,
+            X_test, y_test,
+            doPrint=False,
+            **fit_params
+        )
+
+        accuracies[model_class_name]['cross_val_accuracies'].append(model.get_accuracy_test())
+
+    # train the last time on all the trainset and compute the scores on test set
+    model = model_class()
+    model.fit(
+        train_input, train_target,
+        test_input, test_target,
+        doPrint=False,
+        **fit_params
+    )
+    # model.plot_history()
+    # plt.savefig(model_class_name)
+
+    accuracies[model_class_name]['train_accuracy'] = model.get_accuracy_train()
[…]

### 2019-05-02 main.py

@@ -2,25 +2,12 @@ import torch
 from scripts.data import Data
 from scripts.models_implemented import *
 
-model = 3
+# train the best model
+model = 6
 
 ################### GENERATE DATASETS ###################
 
 d = Data()
-
-train_input = None
-train_target = None
-train_classes = None
-test_input = None
-test_target = None
-test_classes = None
-
-if model == 5 or model == 6: # get 3d images
-    train_input, train_target, train_classes, test_input, test_target, test_classes = d.get_data_3dCNN()
-elif model == 4: #get image in 2d
-    train_input, train_target, train_classes, test_input, test_target, test_classes = d.get_data_2dCNN()
-else: # append img 1 and 2
-    train_input, train_target, train_classes, test_input, test_target, test_classes = d.get_data_flatten()
 
 #########################################################
 
@@ -29,10 +16,10 @@ else: # append img 1 and 2
 
 ###################### 1. LINEAR MODEL #####################
 if(model == 1):
-    # torch.manual_seed(1)
 
     # Train the model
     model_linear = LinearRegression()
+    train_input, train_target, test_input, test_target = LinearRegression.reshape_data(d)
 
     print("Number of parameters: {}".format(model_linear.number_params()))
     print("Number of parameters of feature_extractor: /")
@@ -51,8 +38,8 @@ if(model == 1):
 #################### 2. LOGISTIC MODEL #####################
 elif(model == 2):
     # Train the model
-    # torch.manual_seed(1)
     model_logistic = LogisticRegression()
+    train_input, train_target, test_input, test_target = LogisticRegression.reshape_data(d)
 
     print("Number of parameters: {}".format(model_logistic.number_params()))
     print("Number of parameters of feature_extractor: /")
@@ -71,6 +58,8 @@ elif(model == 2):
 ################## 3. NEURAL NET MODEL #####################
 elif(model == 3):
     model_nn1 = NNModel1Loss()
+    train_input, train_target, test_input, test_target = NNModel1Loss.reshape_data(d)
+
 
     print("Number of parameters: {}".format(model_nn1.number_params()))
     print("Number of parameters of feature_extractor: {}".format(model_nn1.number_params(model_nn1.feature_extractor)))
@@ -89,18 +78,15 @@ elif(model == 3):
 ############## 4. NEURAL NET MODEL(2 LOSSES) ###############
 elif(model == 4):
[…]

### 2019-05-02 scripts/data.py

@@ -13,9 +13,9 @@ class Data(object):
         """
         Flat the input (append images after one other)
         """
-        self.train_input_flatten = self.train_input.clone().reshape(self.train_input.size(0), -1)
-        self.test_input_flatten = self.test_input.clone().reshape(self.test_input.size(0), -1)
-        return self.train_input_flatten, self.test_input_flatten
+        train_input_flatten = self.train_input.clone().type(torch.FloatTensor).reshape(self.train_input.size(0), -1)
+        test_input_flatten = self.test_input.clone().type(torch.FloatTensor).reshape(self.test_input.size(0), -1)
+        return train_input_flatten, test_input_flatten
 
     def get_data(self):
         """
@@ -27,20 +27,46 @@ class Data(object):
         """
         Return data as expected by a 3dCNN layer
         """
-        # unsqueeze: add a dimension, example: (n, d) become (n, d, 1) with unsqueeze(2)
-        return self.train_input.clone().type(torch.FloatTensor).unsqueeze(1), self.train_target.clone().type(torch.FloatTensor).unsqueeze(1), self.train_classes.clone().type(torch.FloatTensor), self.test_input.clone().type(torch.FloatTensor).unsqueeze(1), self.test_target.clone().type(torch.FloatTensor).unsqueeze(1), self.test_classes.clone().type(torch.FloatTensor)
+        train_input, train_target, train_classes, test_input, test_target, test_classes = self.get_data()
+        train_input, train_target, test_input, test_target = train_input.unsqueeze(1), train_target.unsqueeze(1), test_input.unsqueeze(1), test_target.unsqueeze(1)
 
-    def get_data_2dCNN(self):
-        train_input_temp, train_target, train_classes, test_input_temp, test_target, test_classes = self.get_data_3dCNN()
-        return train_input_temp.clone().reshape(train_input_temp.size(0), -1), train_target, train_classes, test_input_temp.clone().reshape(train_input_temp.size(0), -1), test_target, test_classes
+        # unsqueeze: add a dimension, example: (n, d) become (n, d, 1) with unsqueeze(2)
+        return train_input, train_target, test_input, test_target
+
+    def get_data_3dCNN2Loss(self):
+        """
+        Return data as expected by a 3dCNN layer with 2 losses
+        """
+        train_input, train_target, train_classes, test_input, test_target, test_classes = self.get_data()
[…]

### 2019-05-02 scripts/model.py

@@ -55,7 +55,11 @@ class Model(nn.Module):
                 train_target = train_target[indices_shuffled]
 
             # iterate over the batches
+            train_loss = 0
+            train_acc = 0
+            n_batches = 0
             for batch_start in range(0, train_input.shape[0], batch_size):
+                n_batches += 1
                 # get the current batch from the trainset
                 batch_end = batch_start + batch_size
                 train_input_batch = train_input[batch_start:batch_end]
@@ -68,10 +72,15 @@ class Model(nn.Module):
                 self.optimizer.zero_grad()
 
                 # return predicted value (forward method of subclass model)
-                train_loss, train_acc = get_loss_acc(train_input_batch, train_target_batch)
+                train_loss_batch, train_acc_batch = get_loss_acc(train_input_batch, train_target_batch)
+                train_loss += train_loss_batch.item()
+                train_acc += train_acc_batch
 
-                train_loss.backward() # backward propagation of grads
+                train_loss_batch.backward() # backward propagation of grads
                 self.optimizer.step() # update params
+            # do the average across batches
+            train_loss /= n_batches
+            train_acc /= n_batches
 
             # Train loss and accuracy have been computed before the train step, test loss and accuracy after it
             test_acc = None
@@ -80,11 +89,12 @@ class Model(nn.Module):
             # do the same with test if it is passed and overwrite test_acc and test_loss
             if test_input is not None:
                 test_loss, test_acc = get_loss_acc(test_input, test_target)
+                test_loss = test_loss.item()
 
             # add epoch to history saving new values
             self.history.epoch(
-                train_loss=train_loss.item(), train_acc=train_acc,
-                test_loss=test_loss.item(), test_acc=test_acc
+                train_loss=train_loss, train_acc=train_acc,
+                test_loss=test_loss, test_acc=test_acc
             )
 
             # ----- print all the information
@@ -92,6 +102,12 @@ class Model(nn.Module):
                 p()
 
         return self
+
+    def get_accuracy_train(self):
+        return self.history.get_accuracies_train()[-1]
+
+    def get_accuracy_test(self):
+        return self.history.get_accuracies_test()[-1]
 
     def plot_history(self):
         """
@@ -184,6 +200,12 @@ class History():
         Plot accuracies of train and test.
         """
         self.plot_general(self.test_acc, self.train_acc, 'Test accuracy', 'Train accuracy', 'accuracy', True)
+
+    def get_accuracies_train(self):
+        return self.train_acc
+
[…]

### 2019-05-02 scripts/models_implemented.py

@@ -27,6 +27,11 @@ class LinearRegression(Model):
         x = x.view(x.shape[0], -1)
         return self.classifier(x) # return predicted value
 
+    @classmethod
+    def reshape_data(cls, data):
+        return data.get_data_flatten()
+
+
 #################### 2. LOGISTIC MODEL #####################
 class LogisticRegression(Model):
 
@@ -39,7 +44,8 @@ class LogisticRegression(Model):
             'criterion': criterion
         }
         self.classifier = nn.Sequential(
-            nn.Linear(features_in, features_out)
+            nn.Linear(features_in, features_out),
+            nn.Sigmoid()
         )
         self.optimizer = optimizer(self.parameters())
         self.criterion = criterion()
@@ -47,7 +53,11 @@ class LogisticRegression(Model):
     def forward(self, x):
         # flatten the features for the linear layer in the classifier
         x = x.view(x.shape[0], -1)
-        return F.sigmoid(self.classifier(x)) # return predicted value
+        return self.classifier(x) # return predicted value
+
+    @classmethod
+    def reshape_data(cls, data):
+        return data.get_data_flatten()
 
 ################## 3. NEURAL NET MODEL #####################
 class NNModel1Loss(Model):
@@ -86,6 +96,11 @@ class NNModel1Loss(Model):
         # flatten the features for the linear layer in the classifier
         features = features.view(X.shape[0], -1)
         return self.classifier(features)
+
+    @classmethod
+    def reshape_data(cls, data):
+        return data.get_data_flatten()
+
 ############## 4. NEURAL NET MODEL(2 LOSSES) ###############
 class NNModel2Loss(Model):
     def __init__(self,
@@ -165,6 +180,10 @@ class NNModel2Loss(Model):
 
         return bool_pred, digit_pred_left, digit_pred_right
 
+    @classmethod
+    def reshape_data(cls, data):
+        return data.get_data_NN2Loss()
+
 ############ 5. CONVOLUTIONAL NEURAL NETWORK ###############
 # add padding to keep input dimensions
 # a = nn.Conv3d(1, 32, kernel_size=(1, 3, 3), padding=(0, 1, 1))
@@ -228,97 +247,105 @@ class CNNModel1Loss(Model):
         features = features.view(X.shape[0], -1)
         return self.classifier(features)
 
+    @classmethod
+    def reshape_data(cls, data):
+        return data.get_data_3dCNN()
+
[…]

### 2019-05-02 report/paper.tex
remove learning rate from old models + report wip
added plotbox
TODO: fix report (2 losses part), add images of models + add table
write crossvalidation part in conclusion

@@
 \end{itemize}
+Files \texttt{scripts/cross\_validation\_report.py} and \texttt{BoxPlot\_generator.ipynb} are made for report purpose.
 
@@
 The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images (we have used an hot encoding to give a target to each node).
-%check se img1 < img2 o se era l'inverso!
 For these two models the images inputs are passed as 2D images.
@@
 from $0$ to $9$ as value of digit).
-We were expected to have a better result using 2 losses; but since it is a model with just linear layers we could have improved the computation considering both the images independently; we have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
+% NN con 2 losses sembra meglio nel train!!
+% We were expected to have a better result using 2 losses; but since it is a model with just linear layers we could have improved the computation considering both the images independently; we have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
 We have observed that two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
+%TODO: aggiungere 2 img
 
 % Aggiungere un preprocessing step (flatten nel caso delle reti basate sui layer lineari e niente nel caso della rete conv3d)
+
+%UPDATE: 2 losses work better with NN but not with CNN!!!!!!!!!
 
@@
 We have used the dropout to generalize better and then obtain a more robust model.
+%TODO: aggiungere 2 img
 
@@
 we have decided to use a unique seed to compare the models.
+In Table \ref{table:params} we have %something.
 
-% #TODO: (REPORT) for 1 to 10 of seed and to plotbox per far vedere outliers --> pensare al codice
-% #TODO: (REPORT) fare tabella con num\_params per ogni modello
+In the following plotboxs, it is possible to see the accuracies for each models (using cross validation).
+We can notice that the accuracy vary a lot in simple models since they highly depend from weights initialization.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{boxplot.png}
+    \caption{Plotbox that show the accuracy using cross validation. Star symbols represent the average values.}
+    \label{fig:train_acc_box}
+\end{figure}
+
+% Table accuracy, cross validation (numpy - mean +- std), test value
+
+\begin{table}
+\centering
+\caption{Table with number of parameters for each model.}
+\label{table:params}
+\begin{tabular}{|l|l|l|l|l|} 
+\cline{3-5}
+\multicolumn{1}{l}{} &                     & \multicolumn{3}{c|}{\textbf{Parameters}}          \\ 
+\cline{3-5}
+\multicolumn{1}{l}{} &                     & Feature extractor & Classifier & Total   \\ 
+\hline
+                     & Linear Regression   & -                 & 393        & 393     \\ 
+\cline{2-5}
+               & Logistic Regression & -                 & 393        & 393     \\ 
+\cline{2-5}
+ \textbf{Models}   & NN (1 loss)~        & 133504            & 129        & 133633  \\ 
+\cline{2-5}
+                     & NN (2 losses)~      & 133504            & 129        & 136213  \\ 
+\cline{2-5}
+                     & CNN (1 loss)~       & 22650             & 25153      & 47803   \\ 
+\cline{2-5}
+                     & CNN (2 losses)      & 22650             & 25153      & 49773   \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 As future improvement we would like to implement the cross validation.
-%TODO: add
+%TODO: add something
+Then CNN is the best %blabla
 % %%%%%%%%%%%%%%%%%%%%%

### 2019-05-03 report/paper.tex
updated report

--- 2019-04-30 report/paper.tex
@@
 
+%rifare img
 \begin{figure}[!h]
@@
 from $0$ to $9$ as value of digit).
-We were expected to have a better result using 2 losses; but since it is a model with just linear layers we could have improved the computation considering both the images independently; we have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
+We have noticed that NN with 2 losses seems to perform better during the cross validation but not in the test (see Table \ref{table:accuracy}).
+That seems to be in contrast with what we would expected; the cause could be found in the structure of the model, since we just have linear layers we could have improved the computation considering both the images independently.
+We have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
 We have observed that two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
+%TODO: aggiungere 2 img
 
@@
 %AGGIUNGERE COMPARAZIONE PER DIRE QUALE DEI DUE MODELLI É STATO MEGLIO
-We have used the dropout to generalize better and then obtain a more robust model.
+We have used the dropout to generalize better and then obtain a more robust model. CNN with 2 losses seems to return the best accuracy (see Table \ref{table:accuracy}).
+%TODO: aggiungere 2 img
 
@@
 Since linear models are really sensible to the weights setting 
-we have decided to use a unique seed to compare the models.
+we have decided to use a cross validation to compare the models (however, we did not implement our cross validation with \textit{pytorch}, we have used the library \textit{sklearn}).
+Table \ref{table:params} summarizes all the number of parameters for each model.
+In the plotboxs (see Fig.
+\ref{fig:train_acc_box}), it is possible to see the accuracies for each models (using cross validation).
+We can notice that the accuracy vary a lot in simple models since they highly depend from weights initialization.
 
-% #TODO: (REPORT) for 1 to 10 of seed and to plotbox per far vedere outliers --> pensare al codice
-% #TODO: (REPORT) fare tabella con num\_params per ogni modello
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{boxplot.png}
+    \caption{Plotbox that show the accuracy using cross validation. Star symbols represent the average values.}
+    \label{fig:train_acc_box}
+\end{figure}
+
+\begin{table}
+\centering
+\caption{Table with number of parameters for each model.}
+\label{table:params}
+\begin{tabular}{|l|l|l|} 
+\cline{3-3}
+\multicolumn{1}{l}{} &                     & \textbf{Parameters}  \\ 
+\cline{3-3}
+\hline
+                     & Linear Regression  & 393     \\ 
+\cline{2-3}
+               & Logistic Regression & 393     \\ 
+\cline{2-3}
+ \textbf{Models}   & NN (1 loss)~        & 133633  \\ 
+\cline{2-3}
+                     & NN (2 losses)~      &  136213  \\ 
+\cline{2-3}
+                     & CNN (1 loss)~       & 47803   \\ 
+\cline{2-3}
+                     & CNN (2 losses)     & 49773   \\
+\hline
+\end{tabular}
+\end{table}
+
+Table \ref{table:accuracy} shows the comparison between the accuracy obtained during the cross validation and the one of test.
+We can notice that test accuracy got better scores because we have used a bigger dataset to train.
+The dataset used in cross validation was $4/5$ of the total train set, instead the one used to test at the end was $5/5$ (all) of the dataset.
+\begin{table}
+\centering
+\caption{Table accuracy (Cross validation: mean $\pm$ std).}
+\label{table:accuracy}
+\begin{tabular}{|l|l|l|l|} 
+\cline{3-4}
+\multicolumn{1}{l}{} &                     & \multicolumn{2}{c|}{\textbf{Accuracy}}  \\ 
+\cline{3-4}
+\multicolumn{1}{l}{} &                     & Cross validation & Test                 \\ 
+\hline
+                     & Linear Regression   & $61.8\pm7.467$    & $69.2$                 \\ 
+\cline{2-4}
+     & Logistic Regression & $72.4 \pm0.860$    & $73.8$                 \\ 
+\cline{2-4}
+\textbf{Models}    & NN (1 loss)~        & $70.5 \pm12.696$   & $80.5$                 \\ 
+\cline{2-4}
+                     & NN (2 losses)~      & $73.8 \pm10.419$   & $78.4$                 \\ 
+\cline{2-4}
+                     & CNN (1 loss)~       & $81.6 \pm2.289$    & $81.7$                 \\ 
+\cline{2-4}
+                     & CNN (2 losses)      & $81.6 \pm2.223$    & $83.4$                 \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \label{sec:conclusion}
-As future improvement we would like to implement the cross validation.
-%TODO: add
+Then CNN represents the best model with images recognition. In particular, the CNN with 2 losses seems to perform slightly better. 
+%TODO: add something
+As future improvement we would like to implement the cross validation using the \textit{pytorch} library.
 % %%%%%%%%%%%%%%%%%%%%%
@@
 \end{document}
+
+% \begin{table}
+% \centering
+% \caption{Table with number of parameters for each model.}
+% \label{table:params}
+% \begin{tabular}{|l|l|l|l|l|} 
+% \cline{3-5}
+% \multicolumn{1}{l}{} &                     & \multicolumn{3}{c|}{\textbf{Parameters}}          \\ 
+% \cline{3-5}
+% \multicolumn{1}{l}{} &                     & Feature extractor & Classifier & Total   \\ 
+% \hline
+%                      & Linear Regression   & -                 & 393        & 393     \\ 
+% \cline{2-5}
+%               & Logistic Regression & -                 & 393        & 393     \\ 
+% \cline{2-5}
+%  \textbf{Models}   & NN (1 loss)~        & 133504            & 129        & 133633  \\ 
+% \cline{2-5}
+%                      & NN (2 losses)~      & 133504            & 129        & 136213  \\ 
+% \cline{2-5}
+%                      & CNN (1 loss)~       & 22650             & 25153      & 47803   \\ 
+% \cline{2-5}
+%                      & CNN (2 losses)      & 22650             & 25153      & 49773   \\
+% \hline
+% \end{tabular}
+% \end{table}

### 2019-05-03 main.py

@@ -3,7 +3,7 @@ from scripts.data import Data
 from scripts.models_implemented import *
 
 # train the best model
-model = 6
+model = 4
 
 ################### GENERATE DATASETS ###################
 

### 2019-05-04 code/data.py
Fixed a bug in cnn2

@@ -1,4 +1,4 @@
-from scripts.dlc_practical_prologue import generate_pair_sets
+from code.dlc_practical_prologue import generate_pair_sets
 import torch
 
 class Data(object):

### 2019-05-04 code/models_implemented.py

@@ -2,7 +2,7 @@ import torch
 from torch import nn
 from torch import optim
 import torch.nn.functional as F
-from scripts.model import Model
+from code.model import Model
 
 ###################### 1. LINEAR MODEL #####################
 # Give weight to each pixel (this weight is taken from gradient descent)
@@ -297,8 +297,8 @@ class CNNModel2Loss(Model):
 
         # classificatore for digit
         self.classifier_digit = nn.Sequential(
-            # 14 x 14 = 196 since we have flatted
-            nn.Linear(14 * 14, 10),
+            # 8 x 7 x 7 = 392 since we have flatted
+            nn.Linear(8 * 7 * 7, 10),
             nn.Sigmoid()
         )
 
@@ -337,14 +337,14 @@ class CNNModel2Loss(Model):
         if len(x.shape) == 4:
             # Conv3d expects an input of shape (N, C_{in}, D, H, W)
             x = x.unsqueeze(1)
+
         features = self.feature_extractor(x)
+        features_first_image = (features[:, :, 0]).contiguous().view(x.shape[0], -1) #shape = [1000, 8*7*7]
+        features_second_image = (features[:, :, 1]).contiguous().view(x.shape[0], -1)
 
         # flatten for the linear layer in the classifier
-        features = features.view(x.shape[0], -1)
-        first_image = (x[:, :, 0]).view(x.shape[0], -1) #shape = [1000, 14*14]
-        second_image = (x[:, :, 1]).view(x.shape[0], -1)
-
-        return self.classifier_bool(features), self.classifier_digit(first_image), self.classifier_digit(second_image) # return predicted values
+        features_flatten = features.view(x.shape[0], -1)
+        return self.classifier_bool(features_flatten), self.classifier_digit(features_first_image), self.classifier_digit(features_second_image) # return predicted values
 
     @classmethod
     def reshape_data(cls, data):

### 2019-05-04 cross_validation_report.py

@@ -1,6 +1,6 @@
 from sklearn.model_selection import KFold
-from scripts.data import Data
-from scripts.models_implemented import *
+from code.data import Data
+from code.models_implemented import *
 import matplotlib.pylab as plt
 
 # get the data

### 2019-05-04 main.py

@@ -1,9 +1,9 @@
 import torch
-from scripts.data import Data
-from scripts.models_implemented import *
+from code.data import Data
+from code.models_implemented import *
 
 # train the best model
-model = 4
+model = 6
 
 ################### GENERATE DATASETS ###################
 
@@ -118,6 +118,7 @@ elif(model == 5):
 ############ 6. CONVOLUTIONAL NEURAL NETWORK (2 losses) ###############
 else:
     model_cnn2 = CNNModel2Loss()
+
     train_input, train_target, test_input, test_target = CNNModel2Loss.reshape_data(d)
     print("Number of parameters: {}".format(model_cnn2.number_params()))
     print("Number of parameters of feature_extractor: {}".format(model_cnn2.number_params(model_cnn2.feature_extractor)))

### 2019-05-05 report/paper.tex
Updated report and images

@@
 \usepackage{url}        % Written by Donald Arseneau
-%\usepackage{stfloats}  % Written by Sigitas Tolusis
 \usepackage{xcolor}
-%\interdisplaylinepenalty=2500
 \usepackage{listings}
@@
 The task was to compare two digits in a two-channel image.
-Starting from linear model until some more complex non-linear model, we have found out that convolutional neural networks represents the best solution to classify images.
+Starting from linear model until some more complex non-linear model, we have found out that convolutional neural networks represents the best solution to classify images with  around 15\% error rate.
 \end{abstract}
@@
 \label{sec:intro}
-\textit{Convolutional neural network} (CNN) represents the most powerful model to analyse images. Indeed, CNN uses the pixel locality to learn better.
+\textit{Convolutional neural network} (CNN) represents the most powerful model to analyse images because it preserves the dimension of an image, that guarantees a good use of pixel locality to learn better.
+Simple linear models were used at the beginning of the project, in particular \textit{Linear Regression} and \textit{Logistic Regression} (see Section \ref{sec:linearmodel}), we obtained an accuracy of $70\%$ (these models not consider the pixel locality).
+We have then used a more complex model \textit{Neural Network} with one-loss and two-losses (as explained and showed in Section \ref{sec:nnmodel}), we got a better accuracy ($80\%$), but we were just using linear layer.
+Our goal, was to try to take advantage by the locality of the pixel (we want to analyze the context of an image and not just a single value).
+The last model, takes advantage by the context of an image, \textit{Convolutional neural network} we got a very good accuracies ($82-84\%$)  (see Section \ref{sec:cnnmodel}).
 
-We have started the project training simple linear models, in particular \textit{Linear Regression} and \textit{Logistic Regression} (see Section \ref{sec:linearmodel}) but we have notice that these models are too simple and not consider the pixel locality.
-We have then used a more complex model \textit{Neural Network} with one-loss and two-losses (as explained and showed in Section \ref{sec:nnmodel}), we got a bad accuracy caused by the fact that we have a model composed by just linear layer.
-The last model that we have implemented is the \textit{Convolutional neural network}, with this model we got a very good approximation (see Section \ref{sec:cnnmodel}).
-In all the models, we have used \textit{Adam's method} instead of \textit{stochastic gradient descent} as optimizer since it use an adaptive learning rate (the learning rate of SGD has an equivalent type of effect for all the weights/parameters of the model \cite{reference0}).
-As criterion, we have used \textit{Mean Square Error} for all the models.
+In all models, we have used \textit{Adam's method} instead of \textit{stochastic gradient descent} as optimizer since it uses an adaptive learning rate (the learning rate of SGD has an equivalent type of effect for all the weights/parameters of the model \cite{reference0}).
+As criterion, we have used \textit{Mean Square Error}.
 
-For all the models we have explicitely separated in the code the feature extractor, used to process the input and extract high level features, from classifier(s), a simple linear layer that we have used to classify.
+The models implementation consists of an abstract class that acts as super class.
+Then, for each model we have implemented a class and inside them we have explicitly separated the feature extractor, used to process the input and extract high level features, from classifier(s), a simple linear layer that we have used to classify.
 
-Section \ref{sec:data} explain the structure of the dataset, Section \ref{sec:codestruc} describes briefly the content of each file.
-In Section \ref{sec:comparison}, \ref{sec:conclusion} we have showed the performance accuracy obtained by each model with a fixed seed, explaining why we got these results.
+Section \ref{sec:data} explains the structure of the dataset, Section \ref{sec:codestruc} describes briefly the content of each file.
+Section \ref{sec:auxilaryloss} explain which auxiliary loss we have used.
+In Section \ref{sec:comparison} and \ref{sec:conclusion} we have showed the performance accuracy obtained by each model using a cross validation, explaining why we got these results.
 
@@
 \label{sec:codestruc}
-In this section contains the description of each file.
+This section describes the content of each file.
 \begin{itemize}
-    \item \texttt{scripts/data.py}: class to generate the dataset. Contains different methods (e.g. flat the input, get a dataset in 2D or in 3D, enable the hot-encoding).
-    \item \texttt{scripts/model.py}: general class to define a model to train and test it (with corresponding plot and history).
-    \item \texttt{scripts/models\_implemented.py}: contains all the model classes (e.g. NNModel1Loss, CNNModel2Loss).
-    \item \texttt{main.py}: contains examples of each models, in order to call and train a model.
+    \item \texttt{code/data.py}: class to generate the dataset. Contains different methods (e.g. flat the input, get a dataset in 2D or in 3D, enable the hot-encoding).
+    \item \texttt{code/model.py}: general class to define a model to train and test it (with corresponding plot and history).
+    \item \texttt{code/models\_implemented.py}: contains all the model classes (e.g. NNModel1Loss, CNNModel2Loss).
+    \item \texttt{main.py}: contains examples of each model, in order to call and train it.
 \end{itemize}
-Files \texttt{scripts/cross\_validation\_report.py} and \texttt{BoxPlot\_generator.ipynb} are made for report purpose.
+Files \texttt{code/cross\_validation\_report.py} and \texttt{BoxPlot\_generator.ipynb} are made for report purposes.
 
@@
 In the logistic regression we have used a sigmoid as an activation function in the output layer.
-A linear layer accept as input a vector of values, that is the reason why we need to flat the input (the raw input is composed by 2D images).
-In fact a linear layer give a weight to each pixel, then it will accept a vector composed by N rows and D dimensions (columns), in our case it will have size $14 \times 14 \times 2$.
-Linear layers bad generalize images, since it just memorize the train images, that causes a good train accuracy but a very low test accuracy.
+A linear layer accepts as input a vector of values, that is the reason why we need to flat the input (the raw input is composed by 2D images).
+It gives a weight to each pixel, then it will accept a vector composed by N rows and D dimensions (columns), in our case it will have size $2 \times 14 \times 14$.
+Linear layers bad generalize images, since it just memorizes the train images, that causes a good train accuracy but a very low test accuracy.
 
-During the train of linear models we have noticed that sometimes the accuracy and loss could both decreases, that can happen if we have a case like the blue line in Fig. \ref{fig:lossacc}. 
+Linear Regression has a strange behavior in training due to the fact that it has not have a sigmoid and the loss MSE works very bad in a classification task without mapping the value in a range.
+
 \begin{figure}[!h]
     \centering
-    \includegraphics[width=0.4\textwidth]{lossacc.png}
-    \caption{Red line has an higher loss but a better accuracy.
-Instead, blue line has a lower loss but it predicts bad (lower accuracy).
-We can conclude saying that blue line fits well samples that already known.}
-    \label{fig:lossacc}
-\end{figure}
-Linear Regression has a strange behavior due to the fact that it has not have a sigmoid and the loss MSE works very bad in a classification task without mapping the value in a range.
-
-%TODO: mettere size flatted come input???
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.4\textwidth]{linearregression.png}
+    \includegraphics[width=0.35\textwidth]{linearregression.png}
     \caption{Linear Regression model}
@@
 
-%rifare img
 \begin{figure}[!h]
     \centering
-    \includegraphics[width=0.4\textwidth]{logistic.png}
+    \includegraphics[width=0.35\textwidth]{logistic.png}
     \caption{Logistic Regression model}
@@
 \label{sec:nnmodel}
-We have implemented two models for the neural network, the first one with 1 loss and the second one with 2 losses, as showed in Fig.
+We have implemented two models for neural network, the first one with 1 loss and the second one with 2 losses, as showed in Fig.
 \ref{fig:nn1} and Fig.
 \ref{fig:nn2}.
-Both models use the same feature extractors but they differs for the classifier.
-The first one has a boolean classifier that returns 1 if the first image is smaller then the second one.
-The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images (we have used an hot encoding to give a target to each node).
-For these two models the images inputs are passed as 2D images.
+Both models use the same feature extractors but they differs for the classifier(s).
+The first one has a boolean classifier that compares the two digits visible in the two images.
+The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images (class classification: we have used an hot encoding to give a target to each node).
+For these two models the inputs are passed as 2D images. %TODO:sure? o é considerato come vettore?
 Since the feature extractor processes both images together (it does not consider the two images independently), we can use a digit classifier with $20$ nodes as output ($10$ nodes for each images; i.e.
 from $0$ to $9$ as value of digit).
-We have noticed that NN with 2 losses seems to perform better during the cross validation but not in the test (see Table \ref{table:accuracy}).
-That seems to be in contrast with what we would expected; the cause could be found in the structure of the model, since we just have linear layers we could have improved the computation considering both the images independently.
-We have then decided to try with a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
-We have observed that two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
-%TODO: aggiungere 2 img
+We have noticed that NN with 2 losses seems to perform worst during the test then the model with 1 loss (see Table \ref{table:accuracy}).
+That seems to be in contrast with what we would expected; two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
+Since we just have linear layers we could have improved the computation considering both the images independently.
+We have then decided to implement a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.35\textwidth]{nn1.png}
+    \caption{Neural Network model (1 loss)}
+    \label{fig:nn1}
+\end{figure}
 
-% Aggiungere un preprocessing step (flatten nel caso delle reti basate sui layer lineari e niente nel caso della rete conv3d)
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{nn2.png}
+    \caption{Neural Network model (2 losses)}
+    \label{fig:nn2}
+\end{figure}
 
@@
 \label{sec:cnnmodel}
-Two models of convolutional neural networks were trained, both received as input 3D images since we have decided to use a \textit{conv3d}.
-Conv3d are powerful cnn filter that are able to traverse the image in \textit{x}, \textit{y} and \textit{z} axes; that make possible to process both the two images in a independent way.
-CNN are powerfull since they use pixel locality, then instead of flatting the input as done in previous model we can just consider the 3D raw images.
-As done, in Sec.
-\ref{sec:nnmodel} we have decided to compare two models, one with one loss and the other one with two losses, as done with Neural Network models.
-%AGGIUNGERE COMPARAZIONE PER DIRE QUALE DEI DUE MODELLI É STATO MEGLIO
-We have used the dropout to generalize better and then obtain a more robust model. CNN with 2 losses seems to return the best accuracy (see Table \ref{table:accuracy}).
-%TODO: aggiungere 2 img
+Two models of convolutional neural networks were trained, both received an input of 3D images since we have decided to use a \textit{conv3d}.
+Conv3d are powerful cnn filter that are able to traverse the image in \textit{x}, \textit{y} and \textit{z} axes; that make possible to independently process both the two images.
+CNN are powerful since they use pixel locality, instead of flatting the input as done in previous model we consider the two 3D images.
+As done with the model of NN, in Sec.
+\ref{sec:nnmodel}, we have decided to compare two models one with one loss and the other one with two losses (see Fig.
+\ref{fig:cnn1} and Fig.
+\ref{fig:cnn2}).
+We have used the dropout to generalize better and then obtain a more robust model. CNN with 2 losses got the best accuracy with $84\%$  (see Table \ref{table:accuracy}).
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{cnn1.png}
+    \caption{Convolutional Neural Network model (1 loss)}
+    \label{fig:cnn1}
+\end{figure}
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{cnn2.png}
+    \caption{Convolutional Neural Network model (2 losses)}
+    \label{fig:cnn2}
+\end{figure}
+
+\section{Comparison between 1 loss and 2 losses}
+\label{sec:auxilaryloss}
+We have decided to train the Neural Network and Convolutional Neural Network models also with an auxiliary loss.
+In addition of the loss obtained with the boolean classifier, we have added a new classifier that should detect the digit class of both images (the target is then an n-hot-encoding).
+Solving two tasks in parallel can help the model to extract a better representation of the data.
+The use of an auxiliary loss improved of $2\%$ the accuracy in CNN model.
 
@@
 Table \ref{table:params} summarizes all the number of parameters for each model.
-In the plotboxs (see Fig.
-\ref{fig:train_acc_box}), it is possible to see the accuracies for each models (using cross validation).
+As expected neural networks require more parameters to got a good accuracy since they do not have weight sharing.
+Convolutional neural networks use weight sharing (a weight for all the pixels of the same filter) then it requires less parameters to got a good accuracy.
+
+Plotboxs (see Fig.
+\ref{fig:train_acc_box}) show the accuracies for each models (using cross validation).
 We can notice that the accuracy vary a lot in simple models since they highly depend from weights initialization.
@@
 \cline{2-3}
-                     & CNN (2 losses)     & 49773   \\
+                     & CNN (2 losses)     & 51733   \\
 \hline
@@
 
-Table \ref{table:accuracy} shows the comparison between the accuracy obtained during the cross validation and the one of test.
+Table \ref{table:accuracy} shows the comparison between the accuracy obtained during the cross validation and the one in test.
 We can notice that test accuracy got better scores because we have used a bigger dataset to train.
@@
 \cline{2-4}
-                     & CNN (2 losses)      & $81.6 \pm2.223$    & $83.4$                 \\
+                     & CNN (2 losses)      & $81.9 \pm2.538$    & $83.7$                 \\
 \hline
@@
 \label{sec:conclusion}
-Then CNN represents the best model with images recognition. In particular, the CNN with 2 losses seems to perform slightly better. 
-%TODO: add something
+Convolutional Neural Network represents the best model with images recognition.
+In particular, the CNN with 2 losses seems to perform slightly better since it is able to extract better the data.
+We have seen that weight sharing (CNN) improves the accuracy and the robustness of the model, the use of an auxiliary loss in this context improves the results obtained.
+
 As future improvement we would like to implement the cross validation using the \textit{pytorch} library.

### 2019-05-06 report/paper.tex
report finished (project finished!)

@@
 \begin{abstract}
-The aim of this project was to show the impact of weight sharing and the use of auxiliary loss.
-The task was to compare two digits in a two-channel image.
-Starting from linear model until some more complex non-linear model, we have found out that convolutional neural networks represents the best solution to classify images with  around 15\% error rate.
+The aim of this project was to show the impact of weight sharing and the use of an auxiliary loss.
+The task was to compare two digits represented as a two-channel image.
+We have started experimenting with a linear model and, then, we have implemented more complex non-linear and convolutional models.
+Our experiments show that, indeed, the use of weight sharing, i.e.
+convolutional layers, and the use of two losses leads to the best solution in his classification problem, achieving an error rate slightly higher than 15\%.
 \end{abstract}
@@
 \label{sec:intro}
-\textit{Convolutional neural network} (CNN) represents the most powerful model to analyse images because it preserves the dimension of an image, that guarantees a good use of pixel locality to learn better.
-Simple linear models were used at the beginning of the project, in particular \textit{Linear Regression} and \textit{Logistic Regression} (see Section \ref{sec:linearmodel}), we obtained an accuracy of $70\%$ (these models not consider the pixel locality).
-We have then used a more complex model \textit{Neural Network} with one-loss and two-losses (as explained and showed in Section \ref{sec:nnmodel}), we got a better accuracy ($80\%$), but we were just using linear layer.
-Our goal, was to try to take advantage by the locality of the pixel (we want to analyze the context of an image and not just a single value).
-The last model, takes advantage by the context of an image, \textit{Convolutional neural network} we got a very good accuracies ($82-84\%$)  (see Section \ref{sec:cnnmodel}).
+\textit{Convolutional neural network} (CNN) represents the most effective architectures to analyse images because, differently than linear models, they allow to keep the shape of the input samples, which, in turn, by the use of weight sharing, allows to exploit the pixel locality to learn better.
+Simple linear models were used at the beginning of the project, in particular \textit{Linear Regression} and \textit{Logistic Regression} (see Section \ref{sec:linearmodel}), we have obtained an accuracy of around $70\%$ (these models do not consider the pixel locality).
+Then, we have used more complex \textit{Linear Neural Networks} respectively with one-loss and two-losses (as explained and showed in Section \ref{sec:nnmodel}).
+With these we have got a better accuracy ($80\%$), but we were just using linear layers.
+Our goal, was to try to take advantage by the locality of the pixel (we want to analyze the images as whole to exploit the locality of the pixels).
+Therefore, the last model architecture we have tried is a \textit{Convolutional neural network} with which we could improve the accuracy ($82-84\%$) (see Section \ref{sec:cnnmodel}).
 
-In all models, we have used \textit{Adam's method} instead of \textit{stochastic gradient descent} as optimizer since it uses an adaptive learning rate (the learning rate of SGD has an equivalent type of effect for all the weights/parameters of the model \cite{reference0}).
-As criterion, we have used \textit{Mean Square Error}.
-
-The models implementation consists of an abstract class that acts as super class.
-Then, for each model we have implemented a class and inside them we have explicitly separated the feature extractor, used to process the input and extract high level features, from classifier(s), a simple linear layer that we have used to classify.
+In all models, we have used \textit{Adam's method} instead of \textit{Stochastic Gradient Descent (SGD)} as optimizer since it uses a adaptive learning rates, while the SGD has a single, fixed, learning rate for all the weights/parameters of the model \cite{reference0}.
+As criterion, we have used \textit{Mean Square Error} and we also show that this loss can be effectively used in a classification problem only if the output of the model is bounded, e.g.
+by the use of a sigmoid function.
 
@@
 \label{sec:data}
-The data set consists of an input series of $2\times14\times14$ tensor (2 images in grayscale).
-Training and test set is $1000$ pairs each.
-In total we will have six tensors: images ($N\times2\times14\times14$), prediction (boolean), classes of the two digits ($N\times2$), both for training and test sets.
-The MNIST database is composed by handwritten digits (from $0$ to $9$); to load the dataset we have used the provided function \texttt{generate\_pair\_sets(N)}.
+Both the train and the test sets consist of $1000$ samples.
+Each input sample is a tensor of shape $2\times14\times14$ tensor and represents 2 images in grayscale.
+While each target sample is represented by 2 tensors: (1) a single value tensor representing the boolean prediction (greater than), and (2) a two values tensor representing the classes of the two digits.
+The MNIST data set is made of handwritten digits (from $0$ to $9$); to load the dataset we have used the provided function \texttt{generate\_pair\_sets(N)}.
 
@@
 
-\section{Linear Models}
+The models implementation consists of an abstract class that acts as super class and implements all the common functionalities which are shared by the different models, e.g.
+a \texttt{fit()} function.
+Then, we have implemented a class for each model which defines two architectures: (1) a feature extractor, used to process the input and extract high level features, (2) one or more classifier(s) (depending on the number of losses that have been used) that are used to obtain either boolean (greater than) or digit predictions.
+
+\section{Models}
+In this section we explain the various architectures we have implemented.
+For each architecture we provide an image detailing its structure.
+In each image, we differentiate in blue the part that we called "feature extractor" and in red the parts that we treat as classifiers.
+
+\subsection{Linear Models}
 \label{sec:linearmodel}
-We have implemented \textit{Linear Regression} and \textit{Logistic Regression}.
-Regression is represented by a final layer that it is linear.
-In the logistic regression we have used a sigmoid as an activation function in the output layer.
-A linear layer accepts as input a vector of values, that is the reason why we need to flat the input (the raw input is composed by 2D images).
-It gives a weight to each pixel, then it will accept a vector composed by N rows and D dimensions (columns), in our case it will have size $2 \times 14 \times 14$.
-Linear layers bad generalize images, since it just memorizes the train images, that causes a good train accuracy but a very low test accuracy.
+We have implemented \textit{Linear Regression},  showed in Fig.
+\ref{fig:linearregression}, and \textit{Logistic Regression}, showed in Fig.
+\ref{fig:logisticregression}.
+In the logistic regression we have used a sigmoid as an activation function as final layer.
+A linear layer accepts as input a vector of values for each sample, therefore, the input samples have been flattened since our raw input sample are overlaid images.
+Therefore, the linear layer assigns a weight to each pixel.
+In short, in our case, the input of the linear models is a tensor of shape $(N, 2 \times 14 \times 14)=(N, 392)$, where N is the number of samples.
+The risk of using linear layer models is that they bad generalize images and may and up "memorizing" the train images therefore leading to overfitting, i.e.
+good train accuracy but a very low test accuracy.
 
-Linear Regression has a strange behavior in training due to the fact that it has not have a sigmoid and the loss MSE works very bad in a classification task without mapping the value in a range.
+We have observed that Linear Regression has a strange behavior during training due to the fact that it does not have a sigmoid and the loss MSE works very bad in a classification task when the output is unbounded because the MSE can easily end up penalizing a lot also correct predictions.
 
@@
 
-\section{Neural Network Models}
+\subsection{Neural Network Models}
 \label{sec:nnmodel}
@@
 The first one has a boolean classifier that compares the two digits visible in the two images.
-The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images (class classification: we have used an hot encoding to give a target to each node).
-For these two models the inputs are passed as 2D images. %TODO:sure? o é considerato come vettore?
-Since the feature extractor processes both images together (it does not consider the two images independently), we can use a digit classifier with $20$ nodes as output ($10$ nodes for each images; i.e.
+The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images.
+Note that for the digit classification we have used the hot encoding to give a target to each node of he output layer.
+As with the previously explained linear models, the inputs are flattened to obtain a vector of features for each sample.
+Since the feature extractor processes both images together (it does not consider the two images independently), we decided to use a digit classifier with $20$ nodes as output ($10$ nodes for each digit; i.e.
 from $0$ to $9$ as value of digit).
-We have noticed that NN with 2 losses seems to perform worst during the test then the model with 1 loss (see Table \ref{table:accuracy}).
-That seems to be in contrast with what we would expected; two losses could improve the accuracy if the model is more complex, otherwise it would affect negatively the accuracy (as in our case).
-Since we just have linear layers we could have improved the computation considering both the images independently.
-We have then decided to implement a convolutional neural network, since \textit{conv3d} allow to automatically process both the images independently.
+We have noticed that NN with 2 losses seems to perform slightly worse during the test than the model with 1 loss (see Table \ref{table:accuracy}).
+This is in contrast with what we were expecting.
+However, we believe that the use of two losses can improve the performance of the model only if the latter is more powerful enough, otherwise it can negatively affect the accuracy (as in our case).
+We could have improved the computation considering both the images independently, i.e.
+processing them independently but with the same network.
+
 \begin{figure}[!h]
@@
 
-\section{Convolutional Neural Network Models}
+\subsection{Convolutional Neural Network Models}
 \label{sec:cnnmodel}
-Two models of convolutional neural networks were trained, both received an input of 3D images since we have decided to use a \textit{conv3d}.
-Conv3d are powerful cnn filter that are able to traverse the image in \textit{x}, \textit{y} and \textit{z} axes; that make possible to independently process both the two images.
-CNN are powerful since they use pixel locality, instead of flatting the input as done in previous model we consider the two 3D images.
+Finally, we have decided to implement convolutional neural networks.
+CNNs are powerful since they use pixel locality and weight sharing.
+In particular, instead of flattening the input, as done for the previously explained models, we can keep the original shape of the images.
+Moreover, we have decided to employ \textit{conv3d} since they allow to automatically process both the images independently without the need of any preprocessing, i.e.
+the filters not only slide along the x and y axes but also along the z axis (in which we have our two images).
 As done with the model of NN, in Sec.
@@
 
-\section{Comparison between 1 loss and 2 losses}
-\label{sec:auxilaryloss}
-We have decided to train the Neural Network and Convolutional Neural Network models also with an auxiliary loss.
-In addition of the loss obtained with the boolean classifier, we have added a new classifier that should detect the digit class of both images (the target is then an n-hot-encoding).
-Solving two tasks in parallel can help the model to extract a better representation of the data.
-The use of an auxiliary loss improved of $2\%$ the accuracy in CNN model.
-
 \section{Comparison of models}
 \label{sec:comparison}
-Since linear models are really sensible to the weights setting 
-we have decided to use a cross validation to compare the models (however, we did not implement our cross validation with \textit{pytorch}, we have used the library \textit{sklearn}).
+We have decided to use cross validation to obtain a more robust comparison of the models (note that we have implement our cross validation with the library \textit{sklearn}).
 Table \ref{table:params} summarizes all the number of parameters for each model.
-As expected neural networks require more parameters to got a good accuracy since they do not have weight sharing.
-Convolutional neural networks use weight sharing (a weight for all the pixels of the same filter) then it requires less parameters to got a good accuracy.
+As expected neural employ weight sharing.
+Convolutional neural networks use weight sharing (a weight for all the pixels of the same filter) and, therefore, they require a lower number of parameters to get a good accuracy.
 
-Plotboxs (see Fig.
-\ref{fig:train_acc_box}) show the accuracies for each models (using cross validation).
-We can notice that the accuracy vary a lot in simple models since they highly depend from weights initialization.
+Box-plots have been used to show, for each model, the accuracies obtained during the cross validation (see Fig.
+\ref{fig:train_acc_box}).
+We can notice that the accuracy can vary a lot in simple models since they are very sensible to the initialization of the weights.
+Moreover, we observe that the use of an auxiliary loss improved by $2\%$ the accuracy of the CNN model.
 
@@
 Table \ref{table:accuracy} shows the comparison between the accuracy obtained during the cross validation and the one in test.
-We can notice that test accuracy got better scores because we have used a bigger dataset to train.
-The dataset used in cross validation was $4/5$ of the total train set, instead the one used to test at the end was $5/5$ (all) of the dataset.
+We can notice that test accuracies are higher because, while we have used the whole train set to obtain the final models and compute the test accuracy, during the cross validation the train set was divided in 5 splits and only 4 of them have been used for the training.
 \begin{table}
@@
 Convolutional Neural Network represents the best model with images recognition.
-In particular, the CNN with 2 losses seems to perform slightly better since it is able to extract better the data.
-We have seen that weight sharing (CNN) improves the accuracy and the robustness of the model, the use of an auxiliary loss in this context improves the results obtained.
+Moreover, the CNN with 2 losses seems to perform slightly better since it learned to extract high-lever feature representing the two digits.
+We have seen that weight sharing (CNN) improves the accuracy and the robustness of the model, the use of an auxiliary loss in this context improves the obtained results.
 
-As future improvement we would like to implement the cross validation using the \textit{pytorch} library.
+% As future improvement we would like to implement the cross validation using the \textit{pytorch} library.
 % %%%%%%%%%%%%%%%%%%%%%

### 2019-05-12 report/paper.tex
updates

@@
 We have seen that weight sharing (CNN) improves the accuracy and the robustness of the model, the use of an auxiliary loss in this context improves the obtained results.
+Our implementation of CNN takes around $190$ seconds, the architecture that we have implemented requires \texttt{conv3d} and \texttt{batchnorm3d} that maybe are not optimized for CPU.
+In the \texttt{main.py} we implemented the same architecture but using only 2D convolutions, this implementation requires around $160$ seconds.
 

### 2019-05-12 code/data.py

@@ -33,6 +33,16 @@ class Data(object):
         # unsqueeze: add a dimension, example: (n, d) become (n, d, 1) with unsqueeze(2)
         return train_input, train_target, test_input, test_target
 
+    def get_data_2dCNN(self):
+        """
+        Return data as expected by a 3dCNN layer
+        """
+        train_input, train_target, train_classes, test_input, test_target, test_classes = self.get_data()
+        train_input, train_target, test_input, test_target = train_input.unsqueeze(2), train_target.unsqueeze(1), test_input.unsqueeze(2), test_target.unsqueeze(1)
+
+        # unsqueeze: add a dimension, example: (n, d) become (n, d, 1) with unsqueeze(2)
+        return train_input, train_target, test_input, test_target
+
     def get_data_3dCNN2Loss(self):
         """
         Return data as expected by a 3dCNN layer with 2 losses

### 2019-05-12 code/model.py

@@ -37,7 +37,7 @@ class Model(nn.Module):
             doPrint: bool (if true it prints the epochs with loss and accuracy)
         """
         if doPrint:
-            p = Print(self) #self is the model
+            p = Print(self)
 
         def get_loss_acc(input_, target):
             pred = self(input_)

### 2019-05-12 code/models_implemented.py

@@ -251,6 +251,68 @@ class CNNModel1Loss(Model):
     def reshape_data(cls, data):
         return data.get_data_3dCNN()
 
+class CNN2dModel1Loss(Model):
+    """
+    Predicts whether the first image is <= than the second. Only one loss can be applied to the output of this model.
+    Input: (N, 2, 14, 14)
+    Output: (N, 1)
+    """
+    def __init__(self,
+        output_size=1, optimizer = optim.Adam, criterion = nn.MSELoss):
+
+        super(CNN2dModel1Loss, self).__init__()
+
+        self.init_params = {
+            'optimizer': optimizer,
+            'criterion': criterion
+        }
+
+        self.feature_extractor = nn.Sequential(
+            nn.BatchNorm2d(1),
+
+            # padding 2+2 on x-axis, 2+2 on y-axis
+            nn.Conv2d(1, 64, kernel_size=(5, 5), padding=(2, 2)),
+            nn.ReLU(),
+            nn.BatchNorm2d(64),
+
+            nn.Conv2d(64, 32, kernel_size=(3, 3), padding=(1, 1)),
+            nn.ReLU(),
+            nn.Dropout(0.3), # dropout = to make it more general and then have a more robust model
+            nn.BatchNorm2d(32),
+
+            # stride, 1 = filter moves on z-axis (1 pixel), shift by 2 on x-axis, shift by 2 on y-axis
+            # padding 1+1 on x-axis, 1+1 on y-axis
+            nn.Conv2d(32, 8, kernel_size=(3, 3), padding=(1, 1), stride=(2, 2)),
+            nn.ReLU(),
+            nn.Dropout(0.3),
+            nn.BatchNorm2d(8)
+        )
+
+        self.classifier = nn.Sequential(
+            # 2 images
+            # 7 x 7 instead of 14 x 14 as size of the image because we have applied a stride
+            nn.Linear(8 * 2 * 7 * 7, 32),
+            nn.ReLU(),
+            nn.Linear(32, output_size),
+            nn.Sigmoid()
+        )
+
+        self.optimizer = optimizer(self.parameters())
+        self.criterion = criterion()
+
+    def forward(self, X):
+        # image must be in 3d
+        features_l = self.feature_extractor(X[:, 0]).view(X.shape[0], -1)
+        features_r = self.feature_extractor(X[:, 1]).view(X.shape[0], -1)
+
+        # flatten the features for the linear layer in the classifier
[…]

### 2019-05-12 main.py

@@ -2,8 +2,11 @@ import torch
 from code.data import Data
 from code.models_implemented import *
 
+import timeit
+import time
+
 # train the best model
-model = 6
+model = 50
 
 ################### GENERATE DATASETS ###################
 
@@ -116,7 +119,7 @@ elif(model == 5):
     model_cnn1.plot_history()
 
 ############ 6. CONVOLUTIONAL NEURAL NETWORK (2 losses) ###############
-else:
+elif (model == 6):
     model_cnn2 = CNNModel2Loss()
 
     train_input, train_target, test_input, test_target = CNNModel2Loss.reshape_data(d)
@@ -124,11 +127,36 @@ else:
     print("Number of parameters of feature_extractor: {}".format(model_cnn2.number_params(model_cnn2.feature_extractor)))
     print("Number of parameters of classifier: {}".format(model_cnn2.number_params(model_cnn2.classifier_bool)))
 
+    start = time. time()
     model_cnn2.fit(
         train_input, train_target,
         test_input, test_target,
-        epochs=50,
-        doPrint=True
+        epochs=25,
+        batch_size=128,
+        doPrint=False
     )
+    end = time. time()
+    print(end - start)
 
-    model_cnn2.plot_history()
+    # model_cnn2.plot_history()
+
+############ Just for comparison, CNN (1 loss and filter of 2d) ###############
+else:
+    model_cnn2d_1 = CNN2dModel1Loss()
+
+    train_input, train_target, test_input, test_target = CNN2dModel1Loss.reshape_data(d)
+    print("Number of parameters: {}".format(model_cnn2d_1.number_params()))
+    print("Number of parameters of feature_extractor: {}".format(model_cnn2d_1.number_params(model_cnn2d_1.feature_extractor)))
+
+    start = time.time()
+    model_cnn2d_1.fit(
+        train_input, train_target,
+        test_input, test_target,
+        epochs=25,
+        batch_size=128,
+        doPrint=False
+    )
+    end = time.time()
+    print(end - start) #time: 164
+
+    # model_cnn2.plot_history()

### 2019-05-12 code/data.py
data cleaned and commented

@@ -1,31 +1,36 @@
 from code.dlc_practical_prologue import generate_pair_sets
 import torch
 
+"""
+Class to handle the data of MNIST.
+"""
 class Data(object):
+
     def __init__(self, n=1000):
         """
-        Initialize data to train and test.
+        Initialize data to train and test. Call generate_pair_sets(n) of file dlc_practical_prologue.
+        Inputs:
+            - n = number of pairs for training and test sets
+        Outputs:
+            - train_input, test_input: images -> N x 2 x 14 x 14
+            - train_target, test_target: class to predict in range[0,1] -> N
+            - train_classes, test_classes: classes of the two digits in [0, 9] -> N x 2
         """
         super().__init__()
         self.train_input, self.train_target, self.train_classes, self.test_input, self.test_target, self.test_classes = generate_pair_sets(n)
 
-    def flat_input(self):
-        """
-        Flat the input (append images after one other)
-        """
-        train_input_flatten = self.train_input.clone().type(torch.FloatTensor).reshape(self.train_input.size(0), -1)
-        test_input_flatten = self.test_input.clone().type(torch.FloatTensor).reshape(self.test_input.size(0), -1)
-        return train_input_flatten, test_input_flatten
-
     def get_data(self):
         """
-        Return data as float
+        Return a copy of data as float.
         """
         return self.train_input.clone().type(torch.FloatTensor), self.train_target.clone().type(torch.FloatTensor), self.train_classes.clone().type(torch.FloatTensor), self.test_input.clone().type(torch.FloatTensor), self.test_target.clone().type(torch.FloatTensor), self.test_classes.clone().type(torch.FloatTensor)
 
     def get_data_3dCNN(self):
         """
-        Return data as expected by a 3dCNN layer
+        Return data as expected by a 3dCNN layer.
+        Output:
+            - train_input, test_input: images -> N x 1 x 2 x 14 x 14
+            - train_target, test_target: class to predict in range[0,1] -> N x 1
         """
         train_input, train_target, train_classes, test_input, test_target, test_classes = self.get_data()
         train_input, train_target, test_input, test_target = train_input.unsqueeze(1), train_target.unsqueeze(1), test_input.unsqueeze(1), test_target.unsqueeze(1)
@@ -35,45 +40,62 @@ class Data(object):
 
     def get_data_2dCNN(self):
         """
-        Return data as expected by a 3dCNN layer
+        Return data as expected by a 2dCNN layer.
[…]

### 2019-05-12 code/models_implemented.py
added some comments up to nn1

@@ -5,22 +5,26 @@ import torch.nn.functional as F
 from code.model import Model
 
 ###################### 1. LINEAR MODEL #####################
-# Give weight to each pixel (this weight is taken from gradient descent)
+# Give weight to each pixel
 class LinearRegression(Model):
-
+    """
+    A linear layer accepts as input a vector of values for each sample, therefore,
+    the input samples have been flattened since our raw input sample are overlaid images.
+    Therefore, the linear layer assigns a weight to each pixel.
+    """
     def __init__(self, features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
         super(LinearRegression, self).__init__()
-        self.init_params = {
-            'features_in': features_in,
-            'features_out': features_out,
-            'optimizer': optimizer,
-            'criterion': criterion
-        }
+
+        # Applies a linear transformation to the incoming data: y = xw + b
+        # x.shape = [N, 392] | w.shape = [392, 1] | b is a scalar
         self.classifier = nn.Sequential(
             nn.Linear(features_in, features_out)
         )
-        self.optimizer = optimizer(self.parameters())
-        self.criterion = criterion()
+
+        # Adam uses a adaptive learning rates,
+        # SGD has a single, fixed, learning rate for all the weights/parameters of the model.
+        self.optimizer = optimizer(self.parameters()) #Adam
+        self.criterion = criterion() #MSE
 
     def forward(self, x):
         # flatten the features for the linear layer in the classifier
@@ -29,6 +33,12 @@ class LinearRegression(Model):
 
     @classmethod
     def reshape_data(cls, data):
+        """
+        Return data as float with inputs flatten.
+        Output:
+            - train_input, test_input: images -> N x 392
+            - train_target, test_target: class to predict in range[0,1] -> N x 1
+        """
         return data.get_data_flatten()
 
 
@@ -37,16 +47,16 @@ class LogisticRegression(Model):
 
     def __init__(self, features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
         super(LogisticRegression, self).__init__()
-        self.init_params = {
-            'features_in': features_in,
-            'features_out': features_out,
-            'optimizer': optimizer,
-            'criterion': criterion
-        }
+
+        # Applies a linear transformation to the incoming data: y = xw + b
+        # x.shape = [N, 392] | w.shape = [392, 1] | b is a scalar
[…]

### 2019-05-14 code/models_implemented.py
NN with 2 losses

@@ -155,37 +155,47 @@ class NNModel2Loss(Model):
         self.criterion = self.custom_criterion
         self.loss_criterion = criterion()
 
-    #TODO: da qua!!!!!!!!!!
+    # Since we are using an auxiliary loss we need to pass our predicted value as tuple (same for target).
+    # Then we can use MSE on all the elements of tuple.
+    # We give more weight to the loss of boolean classification since we are more interested in comparing 2 digits.
     def custom_criterion(self, train_pred, train_target):
             """
+            Function to compute loss when we have an auxiliary loss.
             Input:
                 - train_pred: tuple of 3 elements
-                - train_target:
+                - train_target: tuple of 3 elements
             """
             bool_pred, digit_pred1, digit_pred2 = train_pred
             bool_target, digit_target1, digit_target2 = train_target
 
             bool_loss = self.loss_criterion(bool_pred, bool_target)
             digit_loss1 = self.loss_criterion(digit_pred1, digit_target1)
-
             digit_loss2 = self.loss_criterion(digit_pred2, digit_target2)
 
-            return 10 * bool_loss + digit_loss1 + digit_loss2 # give more weight to bool_loss (10*bool_loss + 1*digit_loss1 + 1*digit_loss2)
+            return 10 * bool_loss + digit_loss1 + digit_loss2
 
     # predict how accurate predict img1 < img2
     def compute_accuracy(self, y_pred, y_target):
+        """
+            Function to compute accuracy when we have an auxiliary loss.
+            We just compute the accuracy on the boolean prediction.
+            Input:
+                - y_pred: tuple of 3 elements
+                - y_target: tuple of 3 elements
+            """
         bool_pred, digit_pred1, digit_pred2 = y_pred
         bool_target, digit_target1, digit_target2 = y_target
 
         bool_pred = bool_pred.clone()
-        bool_pred[bool_pred>0.5] = 1
-        bool_pred[bool_pred<=0.5] = 0
+        bool_pred[bool_pred > 0.5] = 1
+        bool_pred[bool_pred <= 0.5] = 0
         acc = 100 * ((bool_pred == bool_target).sum().type(torch.FloatTensor).item())
         n = bool_pred.shape[0]
         return acc / n # normalize by divide by length (1000) -> same as mean
 
     def forward(self, X):
-        # X.shape = (N, 2*14*14)
+        # X.shape = (1000, 392) -> (N, 2*14*14)
+
[…]

### 2019-05-14 main.py

@@ -6,7 +6,7 @@ import timeit
 import time
 
 # train the best model
-model = 50
+model = 4
 
 ################### GENERATE DATASETS ###################
 

### 2019-05-14 code/models_implemented.py
CNN 1 loss

@@ -11,6 +11,8 @@ class LinearRegression(Model):
     A linear layer accepts as input a vector of values for each sample, therefore,
     the input samples have been flattened since our raw input sample are overlaid images.
     Therefore, the linear layer assigns a weight to each pixel.
+    Input: (N, 392)
+    Output: (N, 1)
     """
     def __init__(self, features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
         super(LinearRegression, self).__init__()
@@ -44,7 +46,10 @@ class LinearRegression(Model):
 
 #################### 2. LOGISTIC MODEL #####################
 class LogisticRegression(Model):
-
+    """
+    Input: (N, 392)
+    Output: (N, 1)
+    """
     def __init__(self, features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
         super(LogisticRegression, self).__init__()
 
@@ -77,6 +82,10 @@ class LogisticRegression(Model):
 
 ################## 3. NEURAL NET MODEL #####################
 class NNModel1Loss(Model):
+    """
+    Input: (N, 392)
+    Output: (N, 1)
+    """
     def __init__(self,
         features_in=392, features_out=1, optimizer = optim.Adam, criterion = nn.MSELoss):
 
@@ -120,6 +129,10 @@ class NNModel1Loss(Model):
 
 ############## 4. NEURAL NET MODEL(2 LOSSES) ###############
 class NNModel2Loss(Model):
+    """
+    Input: (N, 392)
+    Output: (N, 1)
+    """
     def __init__(self,
         features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
 
@@ -219,38 +232,36 @@ class NNModel2Loss(Model):
         """
         return data.get_data_NN2Loss()
 
-############ 5. CONVOLUTIONAL NEURAL NETWORK ###############
-# add padding to keep input dimensions
-# a = nn.Conv3d(1, 32, kernel_size=(1, 3, 3), padding=(0, 1, 1))
+############ 5. CONVOLUTIONAL NEURAL NETWORK (3D CONV) ###############
 class CNNModel1Loss(Model):
     """
-    Predicts whether the first image is <= than the second. Only one loss can be applied to the output of this model.
-    Input: (N, 2, 14, 14)
+    Predicts whether the first image is <= than the second.
+    Only one loss can be applied to the output of this model.
[…]

### 2019-05-14 code/models_implemented.py
CNN 2 losses

@@ -208,7 +208,6 @@ class NNModel2Loss(Model):
 
     def forward(self, X):
         # X.shape = (1000, 392) -> (N, 2*14*14)
-
         # extract features
         features = self.feature_extractor(X)
 
@@ -372,22 +371,23 @@ class CNN2dModel1Loss(Model):
         return data.get_data_2dCNN()
 
 ############ 6. CONVOLUTIONAL NEURAL NETWORK (2 losses) ###############
-# Output = 11 (10 possible values (0 to 9) + 1 to check if image_1 <= image_2)
 class CNNModel2Loss(Model):
-
+    """
+    Two losses can be applied to the output of this model.
+    Input: (N, 1, 2, 14, 14)
+    Output: (N, 11) -> (10 possible values (0 to 9) + 1 to check if image_1 <= image_2)
+    """
     def __init__(self, features_in=392, output_size=11, optimizer=optim.Adam, criterion=nn.MSELoss):
 
         super(CNNModel2Loss, self).__init__()
 
-        self.init_params = {
-            'optimizer': optimizer,
-            'criterion': criterion
-        }
-
         self.feature_extractor = nn.Sequential(
+            # Use batch normalization to normalize the input layer by adjusting and scaling the activations.
+            # Allow each layer of a network to learn by itself a little bit more independently of other layers.
             nn.BatchNorm3d(1),
 
-            # padding 2+2 on x-axis, 2+2 on y-axis
+            # padding 2+2 on x-axis, 2+2 on y-axis (depth=0, height=2, width=2)
+            # add padding to keep input dimensions
             nn.Conv3d(1, 64, kernel_size=(1, 5, 5), padding=(0, 2, 2)),
             nn.ReLU(),
             nn.BatchNorm3d(64),
@@ -397,6 +397,7 @@ class CNNModel2Loss(Model):
             nn.Dropout(0.3), # dropout = to make it more general and then have a more robust model
             nn.BatchNorm3d(32),
 
+            # Stride controls how the filter convolves around the input volume. -> it is used to avoid to have a fraction in output volume instead of an integer
             # stride, 1 = filter moves on z-axis (1 pixel), shift by 2 on x-axis, shift by 2 on y-axis
             # padding 1+1 on x-axis, 1+1 on y-axis
             nn.Conv3d(32, 8, kernel_size=(1, 3, 3), padding=(0, 1, 1), stride=(1, 2, 2)),
@@ -417,7 +418,7 @@ class CNNModel2Loss(Model):
 
         # classificatore for digit
[…]

### 2019-05-14 code/model.py
model commented and cleaned

@@ -5,21 +5,16 @@ import torch
 from torch import nn
 
 from matplotlib.pylab import plt
-from matplotlib.ticker import MaxNLocator
 
 ##################### MODEL CLASS #####################
-
 class Model(nn.Module):
     """
     Abstract class to implement model.
-    Subclass of this module should implement the following methods:
-    - self.init_params: a map composed by
-        - self.optim: pointer to optimizer passed in init.
-        - self.criterion: pointer to loss fn passed in init.
     """
+
     def __init__(self):
         """
-        Initialize function.
+        Initialize function for model and history.
         """
         super(Model, self).__init__()
         self.history = History()
@@ -32,27 +27,35 @@ class Model(nn.Module):
             train_target: train target data
             test_input: test data if any
             test_target: test target data if any
-            epochs: int,
-            batch_size: int,
+            epochs: number of epochs,
+            batch_size: size of the batch,
             doPrint: bool (if true it prints the epochs with loss and accuracy)
         """
+
+        # Iniziale Print to print accuracy and loss for each epoch
         if doPrint:
             p = Print(self)
 
-        def get_loss_acc(input_, target):
-            pred = self(input_)
-            loss = self.criterion(pred, target) # apply loss (i.e. MSE)
-            acc = self.compute_accuracy(pred, target) # compute accuracy
-            return loss, acc
+        def get_loss_accuracy(input_, target):
+            """
+            Function to get the loss and accuracy. Given the input and target.
+            """
+            predicted = self(input_) # get predicted (forward method in each subclass model)
+            loss = self.criterion(predicted, target) # apply loss (i.e. MSE)
+            accuracy = self.compute_accuracy(predicted, target) # compute accuracy
+            return loss, accuracy
 
         for e in range(1, epochs + 1):
+
             # shuffle the train set to select different batches at each epoch
-            indices_shuffled = torch.randperm(train_input.shape[0])
+            indices_shuffled = torch.randperm(train_input.shape[0]) # random permutation of integers from 0 to train_input.shape[0] - 1.
             train_input = train_input[indices_shuffled]
-            if isinstance(train_target, tuple): #bool, digit1, digit2 -> take corresponding in each target
+
+            # shuffle target set according to indices_shuffled
[…]

### 2019-05-14 code/models_implemented.py
cross_validation

@@ -375,9 +375,9 @@ class CNNModel2Loss(Model):
     """
     Two losses can be applied to the output of this model.
     Input: (N, 1, 2, 14, 14)
-    Output: (N, 11) -> (10 possible values (0 to 9) + 1 to check if image_1 <= image_2)
+    Output: (N, 21) -> (10 possible values (0 to 9) for each image + 1 to check if image_1 <= image_2)
     """
-    def __init__(self, features_in=392, output_size=11, optimizer=optim.Adam, criterion=nn.MSELoss):
+    def __init__(self, features_in=392, output_size=21, optimizer=optim.Adam, criterion=nn.MSELoss):
 
         super(CNNModel2Loss, self).__init__()
 

### 2019-05-14 cross_validation_report.py

@@ -3,6 +3,10 @@ from code.data import Data
 from code.models_implemented import *
 import matplotlib.pylab as plt
 
+"""
+This file use library that are not allowed for this project such as sklearn. We have used this file just to do a cross validation
+in order to show results in our report.
+"""
 # get the data
 d = Data()
 
@@ -10,6 +14,7 @@ d = Data()
 n_splits=5
 
 # define the parameters for the fit of the models
+# some models require less epochs (we pick the values looking from the plot)
 model_class_fit_params = {
     LinearRegression: {'epochs': 1000, 'batch_size': 1000},
     LogisticRegression: {'epochs': 1000, 'batch_size': 1000},
@@ -28,16 +33,19 @@ for model_class, fit_params in model_class_fit_params.items():
     train_input, train_target, test_input, test_target = model_class.reshape_data(d)
 
     accuracies[model_class_name] = {'cross_val_accuracies': [], 'train_accuracy': None, 'test_accuracy': None}
-    # cross validdation estimate the score we will obtain with the test set
+
+    # cross validation estimate the score we will obtain with the test set
     for train_index, test_index in kf.split(train_input):
         X_train, X_test = train_input[train_index], train_input[test_index]
-        if isinstance(train_target, tuple):
+
+        if isinstance(train_target, tuple): #use auxiliary loss
             y_train = tuple([y[train_index] for y in train_target])
             y_test = tuple([y[test_index] for y in train_target])
         else:
             y_train, y_test = train_target[train_index], train_target[test_index]
 
         model = model_class()
+
         model.fit(
             X_train, y_train,
             X_test, y_test,
@@ -55,6 +63,7 @@ for model_class, fit_params in model_class_fit_params.items():
         doPrint=False,
         **fit_params
     )
+
     # model.plot_history()
     # plt.savefig(model_class_name)
 

### 2019-05-14 main.py

@@ -6,16 +6,11 @@ import timeit
 import time
 
 # train the best model
-model = 4
+model = 6
 
 ################### GENERATE DATASETS ###################
 
 d = Data()
-
-#########################################################
-
-# Adam’s method considered as a method of Stochastic Optimization is a technique implementing adaptive learning rate.
-# Whereas in normal SGD the learning rate has an equivalent type of effect for all the weights/parameters of the model.
 
 ###################### 1. LINEAR MODEL #####################
 if(model == 1):
@@ -133,12 +128,12 @@ elif (model == 6):
         test_input, test_target,
         epochs=25,
         batch_size=128,
-        doPrint=False
+        doPrint=True
     )
     end = time. time()
     print(end - start)
 
-    # model_cnn2.plot_history()
+    model_cnn2.plot_history()
 
 ############ Just for comparison, CNN (1 loss and filter of 2d) ###############
 else:
@@ -154,9 +149,9 @@ else:
         test_input, test_target,
         epochs=25,
         batch_size=128,
-        doPrint=False
+        doPrint=True
     )
     end = time.time()
     print(end - start) #time: 164
 
-    # model_cnn2.plot_history()
+    model_cnn2d_1.plot_history()

### 2019-05-14 report/paper.tex
updated report

@@
 The second model, it has also a digit classifier that returns the corresponding digit for both the handwritten images.
-Note that for the digit classification we have used the hot encoding to give a target to each node of he output layer.
+Note that for the digit classification we have used the hot encoding to give a target to each node of the output layer.
 As with the previously explained linear models, the inputs are flattened to obtain a vector of features for each sample.
@@
 from $0$ to $9$ as value of digit).
-We have noticed that NN with 2 losses seems to perform slightly worse during the test than the model with 1 loss (see Table \ref{table:accuracy}).
+We have noticed that NN with 2 losses seems to perform slightly worst during the test than the model with 1 loss (see Table \ref{table:accuracy}).
 This is in contrast with what we were expecting.

### 2019-05-14 code/data.py
final changes

@@ -1,4 +1,4 @@
-from code.dlc_practical_prologue import generate_pair_sets
+from dlc_practical_prologue import generate_pair_sets
 import torch
 
 """

### 2019-05-14 code/models_implemented.py

@@ -131,10 +131,10 @@ class NNModel1Loss(Model):
 class NNModel2Loss(Model):
     """
     Input: (N, 392)
-    Output: (N, 1)
+    Output: (N, 21)
     """
     def __init__(self,
-        features_in=392, features_out=1, optimizer=optim.Adam, criterion=nn.MSELoss):
+        features_in=392, features_out=21, optimizer=optim.Adam, criterion=nn.MSELoss):
 
         super(NNModel2Loss, self).__init__()
 

### 2019-05-14 main.py

@@ -26,7 +26,7 @@ if(model == 1):
     model_linear.fit(
         train_input, train_target,
         test_input, test_target,
-        epochs=50,
+        epochs=1000,
         batch_size=1000, #so fast that batch does not make sense to use batch
         doPrint=True
     )
@@ -66,7 +66,7 @@ elif(model == 3):
     model_nn1.fit(
         train_input, train_target,
         test_input, test_target,
-        epochs=50,
+        epochs=300,
         batch_size=1000, #so fast that batch does not make sense to use batch
         doPrint=True
     )
@@ -85,7 +85,7 @@ elif(model == 4):
     model_nn2.fit(
         train_input, train_target,
         test_input, test_target,
-        epochs=50,
+        epochs=300,
         batch_size=1000, #so fast that batch does not make sense to use batch
         doPrint=True
     )
@@ -107,7 +107,8 @@ elif(model == 5):
     model_cnn1.fit(
         train_input, train_target,
         test_input, test_target,
-        epochs=50,
+        epochs=25,
+        batch_size=128,
         doPrint=True
     )
 

### 2019-05-14 main.py
added readme

@@ -94,10 +94,6 @@ elif(model == 4):
 
 ############ 5. CONVOLUTIONAL NEURAL NETWORK ###############
 elif(model == 5):
-
-    # add another dimension (as expected from Conv3d)
-    # X = train_input.unsqueeze(1)
-    # a(X).shape --> 32 filters => in the second dimensions we have 32 "layers" and each one is calculated by a filter
     model_cnn1 = CNNModel1Loss()
     train_input, train_target, test_input, test_target = CNNModel1Loss.reshape_data(d)
     print("Number of parameters: {}".format(model_cnn1.number_params()))
@@ -123,7 +119,6 @@ elif (model == 6):
     print("Number of parameters of feature_extractor: {}".format(model_cnn2.number_params(model_cnn2.feature_extractor)))
     print("Number of parameters of classifier: {}".format(model_cnn2.number_params(model_cnn2.classifier_bool)))
 
-    start = time. time()
     model_cnn2.fit(
         train_input, train_target,
         test_input, test_target,
@@ -131,12 +126,10 @@ elif (model == 6):
         batch_size=128,
         doPrint=True
     )
-    end = time. time()
-    print(end - start)
 
     model_cnn2.plot_history()
 
-############ Just for comparison, CNN (1 loss and filter of 2d) ###############
+############ Just for comparison, CNN (1 loss and 2d conv) ###############
 else:
     model_cnn2d_1 = CNN2dModel1Loss()
 
@@ -144,7 +137,7 @@ else:
     print("Number of parameters: {}".format(model_cnn2d_1.number_params()))
     print("Number of parameters of feature_extractor: {}".format(model_cnn2d_1.number_params(model_cnn2d_1.feature_extractor)))
 
-    start = time.time()
+    # start = time.time()
     model_cnn2d_1.fit(
         train_input, train_target,
         test_input, test_target,
@@ -152,7 +145,7 @@ else:
         batch_size=128,
         doPrint=True
     )
-    end = time.time()
-    print(end - start) #time: 164
+    # end = time.time()
+    # print(end - start) #time: 164
 
     model_cnn2d_1.plot_history()

### 2019-05-16 report-latex/paper.tex
added latex report

--- 2019-05-14 report/paper.tex
@@
 Our experiments show that, indeed, the use of weight sharing, i.e.
-convolutional layers, and the use of two losses leads to the best solution in his classification problem, achieving an error rate slightly higher than 15\%.
+convolutional layers, and the use of two losses leads to the best solution in his classification problem, achieving an error rate around 15\%.
 \end{abstract}
@@
 Section \ref{sec:data} explains the structure of the dataset, Section \ref{sec:codestruc} describes briefly the content of each file.
-Section \ref{sec:auxilaryloss} explain which auxiliary loss we have used.
 In Section \ref{sec:comparison} and \ref{sec:conclusion} we have showed the performance accuracy obtained by each model using a cross validation, explaining why we got these results.
@@
 While each target sample is represented by 2 tensors: (1) a single value tensor representing the boolean prediction (greater than), and (2) a two values tensor representing the classes of the two digits.
-The MNIST data set is made of handwritten digits (from $0$ to $9$); to load the dataset we have used the provided function \texttt{generate\_pair\_sets(N)}.
+The MNIST data set is made of handwritten digits (from $0$ to $9$); to load the dataset we have used the provided function \texttt{generate\_pair\_sets(N)} in file \texttt{dlc\_practical\_prologue.py}.
 
@@
 \end{itemize}
-Files \texttt{code/cross\_validation\_report.py} and \texttt{BoxPlot\_generator.ipynb} are made for report purposes.
+Files \texttt{cross\_validation\_report.py} and \texttt{BoxPlot\_generator.ipynb} are made for report purposes.
 
@@
 Our implementation of CNN takes around $190$ seconds, the architecture that we have implemented requires \texttt{conv3d} and \texttt{batchnorm3d} that maybe are not optimized for CPU.
-In the \texttt{main.py} we implemented the same architecture but using only 2D convolutions, this implementation requires around $160$ seconds.
+In the \texttt{models\_implemented.py} we implemented the same architecture but using only 2D convolutions, this implementation requires around $160$ seconds.
 
